
---
title: "github_document"
author: "Maggie Hsu"
output: html_document
date: "2024-09-12"
---

# Problem 1
```{r problem 1 load dataset}
#Load the penguins dataset from the palmerpenguins package
data("penguins", package = "palmerpenguins")
```
## Data Description
The dataset for this problem is "Penguins" from the "Palmer Penguins" package, with the variables `r names(penguins) `. There are `r nrow(penguins)` rows and `r ncol(penguins)` columns within the dataset. 

Among the penguins in the dataset with known flipper lengths, the mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` millimeters.

## Scatterplot: Bill Length vs. Flipper Length
```{r problem 1 scatterplot, warning=FALSE}
#Create a scatterplot comparing bill length and flipper length in millimeters.
library(ggplot2)
scatter_bill_flipper <- ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm, color=species))+geom_point()
scatter_bill_flipper
ggsave("penguin.png", scatter_bill_flipper) #Save and export the scatterplot as a PNG
```