
---
title: "github_document"
author: "Maggie Hsu"
output: html_document
date: "2024-09-12"
---

# Problem 1
```{r problem 1 load dataset}
#Load the penguins dataset from the palmerpenguins package
data("penguins", package = "palmerpenguins")

#Load the tidyverse package for analyzing data
library("tidyverse")

```
## Data Description
The dataset for this problem is "Penguins" from the "Palmer Penguins" package, with the variables `r names(penguins) `. There are `r nrow(penguins)` rows and `r ncol(penguins)` columns within the dataset. 

Among the penguins in the dataset with known flipper lengths, the mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` millimeters.

## Scatterplot: Bill Length vs. Flipper Length
```{r problem 1 scatterplot, warning=FALSE}
#Create a scatterplot comparing bill length and flipper length in millimeters.
scatter_bill_flipper <- ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm, color=species))+geom_point()

#Display and save the penguin scatterplot. 
scatter_bill_flipper
ggsave("penguin.png", scatter_bill_flipper) #Save and export the scatterplot as a PNG
```

# Problem 2
```{r problem 2 data frame}
## Set up a data frame consisting of the specified vectors
dataframe_problem2 <- tibble(
  #Generate vectors within the data frame
  normal_sample = rnorm(n=10), #A vector consisting of a random sample from a standard normal distribution
  greater_than_0 = normal_sample > 0, #A logical vector indicating if a given value in the sample is greater than 0
  characters = c("a","b","c","d","e","f","g","h","i","j"), #A character vector with length 10
  levels = as.factor(c(1,0,2,1,2,0,1,2,2,0)) #A vector consisting of factors with length 10 and 3 levels
)

#Calculate the means for each variable within the dataframe
mean(pull(dataframe_problem2, var="normal_sample")) #Mean of the "normal_sample" vector
mean(pull(dataframe_problem2, var="greater_than_0")) #Mean for the "greater_than_0" vector
mean(pull(dataframe_problem2, var="characters")) #Mean for the "characters" vector
mean(pull(dataframe_problem2, var="levels")) #Mean for the "levels" vector
```
The means for the "normal sample" and "greater than 0" vectors both result in calculated, numeric values. However, the mean() command returns NA for the characters and level vectors since they do not contain numeric or logical variables.

```{r problem 2 numeric conversion}
#Convert the logical, character, and factor vectors into numeric vectors using the as.numeric function
num_greater_than_0 <- as.numeric(greater_than_0)
num_characters <- as.numeric(characters)
num_levels <- as.numeric(levels)

#Find the mean of each vector's variables.
mean_num_greater <- mean(num_greater_than_0)
mean_num_chars <- mean(num_characters)
mean_num_levels <- mean(num_levels)
```

```{r problem 2 converted means, echo=FALSE}
#Print and display means after converting vectors to numeric types.
mean_num_greater
mean_num_chars 
mean_num_levels 
```
For the "greater than" and level vectors, conversion with as.numeric() results in numeric variables and therefore a numeric mean. However, the character vector was unable to be converted using as.numeric(), resulting in an NA output after the mean function is applied. This explains the mean() command's outputs only working for numeric vectors and vectors with variables that can be converted or coerced into numeric types. 