
---
title: "github_document"
author: "Maggie Hsu"
output: html_document
date: "2024-09-12"
---

# Problem 1
```{r problem 1 load dataset}
#Load the penguins dataset from the palmerpenguins package
data("penguins", package = "palmerpenguins")
```
## Data Description
The dataset for this problem is "Penguins" from the "Palmer Penguins" package, with the variables `r names(penguins) `. There are `r nrow(penguins)` rows and `r ncol(penguins)` columns within the dataset. 

Among the penguins in the dataset with known flipper lengths, the mean flipper length is `r mean(penguins$flipper_length_mm, na.rm = TRUE)` millimeters.

## Scatterplot: Bill Length vs. Flipper Length
```{r problem 1 scatterplot, warning=FALSE}
#Create a scatterplot comparing bill length and flipper length in millimeters.
library(ggplot2)
scatter_bill_flipper <- ggplot(penguins, aes(x=bill_length_mm, y=flipper_length_mm, color=species))+geom_point()
scatter_bill_flipper
ggsave("penguin.png", scatter_bill_flipper) #Save and export the scatterplot as a PNG
```

# Problem 2
```{r problem 2 data frame}
#Generate vectors
normal_sample <- rnorm(10) #random sample from standard normal distribution
greater_than_0 <- normal_sample > 0 #logical vector indicating if a given value in the sample is greater than 0
characters <- c("a","b","c","d","e","f","g","h","i","j") #character vector with length 10
levels <- as.factor(c(1,0,2,1,2,0,1,2,2,0)) #vector consisting of factors with length 10 and 3 levels

dataframe_problem2 <- data.frame(normal_sample,greater_than_0,characters,levels)

#Find the means of each vector
mean_normal_sample <- mean(normal_sample)
mean_greater <- mean(greater_than_0)
mean_chars <- mean(characters)
mean_levels <- mean(levels)
```
The means for the "normal sample" and "greater than 0" vectors both result in calculated, numeric values. However, the mean() command returns NA for characters and levels since these vectors do not contain numeric or logical variables.

```{r problem 2 numeric conversion}
as.numeric
```